---
title: 'Homework 9 - statistical analysis of real and randomized data'
author: "Morgan Southgate"
date: "March 29, 2017"
output:
  html_document: default
  pdf_document: default
---

## Import Adiantum quantitative ecological data
```{r}
adDat <- read.table("QuantitativeDataAdiantum1.csv",sep=",",header=T,stringsAsFactors = F)
head(adDat,3)
tail(adDat,3)

# Clean up data set by changing NA values to 0 - justified for this data-set where the missing values are an absent soil layer and no test results for heavy metal presence from rich northern hardwood forest sites
# Transform NA values to 0
adDat[is.na(adDat)]<- 0
head(adDat,3)
dim(adDat)
```

# Part 1: Statisticlly analyze real data

## Linear Regression of Ca:Mg ratio in soil
### Serpentine soil is characterized by the calcium to magnesium ratio. Are these two elements inversely related to one another in the dataset that I gathered?
```{r, include=FALSE}
# Put functions in here
##Linear regression function
linReg <- function(xVar= 1:100,yVar=runif(100,min=0,max=100)){
  dataframe=data.frame(xVar,yVar)
  linRegMod <- lm(yVar~xVar,data=dataframe)
  linRegOut <- c(slope=summary(linRegMod)$coefficients[2,1],
                 pValue=summary(linRegMod)$coefficients[2,4])
    return(linRegOut)}

linRegPlot <-  function(xVar=1:100,yVar=runif(100,min=0,max=100)){
  dataframe=data.frame(xVar,yVar)
  lrplot <- plot(y=dataframe$yVar,
                       x=dataframe$xVar,
                      cex=2, 
                      pch=21,
                      bg="purple",
                      main="Linear Regression",
                      xlab="X Variable", 
                      ylab="Y variable",
                      xlim=range(xVar),
                      ylim=range(yVar))
  linRegMod <- lm(yVar~xVar,data=dataframe)
  abline(linRegMod)
  return(lrplot)
}

```


```{r, echo=T}
# Use the linear regression function to test the relationship between soil concentrations of Ca and Mg in adDat
linReg(xVar=adDat$Ca,yVar=adDat$Mg)

# plot results using linRegPlot
linRegPlot(xVar=adDat$Ca,yVar=adDat$Mg)

# Based on this analysis the concentrations of Ca and Mg in my soil sample data appear to have a significant inverse relationship
```

## Logistic Regression of Depth A Soil Layer between the serpentine species *A. aleuticum* & *A. viridimontanum*

```{r, include=FALSE}
# Function for logistic regression in R from HW 8
logReg <- function(xVar=rgamma(n=20,shape=5,scale=5), 
                   yVar=rbinom(n=20,size=1,p=0.5)){
  logRegMod <- glm(yVar~ xVar,
                   family=binomial(link="logit"))
  logRegOut <- c(xVarEst=summary(logRegMod)$coefficients[2,1],
                 pValue=summary(logRegMod)$coefficients[2,4])
   return(logRegOut)}

# Function for plotting logistic regression results in R from HW8
logRegPlot <- function(xVar=rgamma(n=20,shape=5,scale=5), 
                       yVar=rbinom(n=20,size=1,p=0.5),dataframe = data.frame(xVar,yVar)){
lrResults = glm(yVar ~ xVar, family="binomial"(link="logit"))
LRplot = plot(dataframe$xVar, y=dataframe$yVar,
                   pch=21,
                   bg="tan",
                   cex=1,
                   main="Logistic Regression")
LRplot = curve(predict(lrResults,
              data.frame(xVar=x),
              type="response"),
      add=TRUE,
      lwd=2)
return(LRplot)}

logRegPlot()
```

```{r, echo=TRUE}
# subset out just the serpentine maidenhair data from adDat
serpDatA <- subset(adDat,SpeciesID=="A")
serpDatV <- subset(adDat,SpeciesID=="V")

# create new species id column to be used in place of letter species id for logistic regression, so that A. aleuticum species id=0 and A. viridimontanum species id=1
dim(serpDatA)
dim(serpDatV)

id <- rep("0",length=6)
serpDatA <- cbind(serpDatA,id)

id <- rep("1",length=9)
serpDatV <- cbind(serpDatV,id)

# combine both subsets into a single data frame serpDat
serpDat <- rbind(serpDatA,serpDatV)

# Test the difference in depth of the A soil layer between the two serpentine maidenhairs using logistic regression - the difference is non-significant, with a p-value of .13. 
logReg(xVar=serpDat$AvgDepthA,yVar=serpDat$id)
logRegPlot(xVar=serpDat$AvgDepthA,yVar=serpDat$id)

```


##Test variation in depth of the A layer between species groups using ANOVA

```{r, include=FALSE}
ANOV <- function(xVar=as.factor(rep(c("A","B","C","D","E"),each=3)),
                    yVar=c(rgamma(10,shape=5,scale=5),rgamma(5,shape=5,scale=10))){
                      df=data.frame(xVar,yVar)
          aovMod <- aov(yVar~xVar,data=df)
          aovOut <- summary(aovMod)[[1]][["Pr(>F)"]][1]
         return(aovOut)}

ANOVplot <- function(xVar=as.factor(rep(c("A","B","C","D","E"),each=3)),yVar=c(rgamma(15,shape=5,scale=5))){
  df <- data.frame(xVar,yVar)
  aovMod <- aov(yVar~xVar,data=df)
  aovPlot <-  boxplot(yVar~xVar,
                      data=df,
                      col=c("red","yellow","blue","green","purple"),
                    xlab=names(xVar),ylab=names(yVar))
  return(aovPlot)}
  
ANOVplot()
```


```{r}
# run anova on the avg depth of the A layer of the soil 
ANOV(xVar=adDat$SpeciesID,yVar=adDat$AvgDepthA)
ANOVplot(xVar=adDat$SpeciesID,yVar=adDat$AvgDepthA)

```

# Part 2: Statistical Analysis of Randomized Data

## Make a function ranDat that measures the parameters of a vector or a column of a dataframe using the fitdistr function from the MASS package and the rgamma function to simulate continuous random values over 0.
```{r}
library(MASS)
ranDat <- function(var=rnorm(29,10,1),dist="gamma",n=29){
  fitvar =  fitdistr(var,densfun=dist)
  ranvar = rgamma(n,shape=fitvar$estimate[1],scale=fitvar$estimate[2])
                return(print(ranvar))}

ranDat()
```

# Use the ranDat() function to generate mock data of the Ca and Mg levels in the soil with distributions based on the parameters of the original data. 
```{r}
library(MASS)
ranMg <- ranDat(adDat$Mg)
ranCa <- ranDat(adDat$Ca)

# bind into a single data frame
randomData <- data.frame(ranCa,ranMg)
head(randomData)

# test the relationship between the randomized Ca and Mg concentrations using the linear regression function
linReg(xVar=randomData$ranCa,yVar=randomData$ranMg)
linRegPlot(xVar=randomData$ranCa,yVar=randomData$ranMg)

```

## Test differences in the percent organic matter of the soil between the two species of serpentine maidenhair ferns with ANOVA using randomized data
```{r}
# Subset out percent organic matter for just the serpentine maidenhairs
serpA <- subset(adDat,SpeciesID=="A")
serpV <- subset(adDat,SpeciesID=="V")

serpSpp <- rbind(serpA,serpV)

# use the fitdistr function to get parameters of percent organic matter in soil using gamma distribution
ranOM <- ranDat(serpSpp$Ompct,"gamma",n=15)

# Create vector of species id's and assign to randomly generated ranOM values
ranSpID <- rbinom(15,size=1,prob=.5)
randomData1 <- data.frame(ranSpID,ranOM)

#Run ANOVA with randomly generated values - returns non-significant result
ANOV(xVar=randomData1$ranSpID,yVar=randomData1$ranOM)
ANOVplot(xVar=randomData1$ranSpID,yVar=randomData1$ranOM)
```

# Part 3: Statistical Analysis of Idealized (*bling*) data

## Modeling a significant difference in soil percent organic matter between the two serpentine maidenhair ferns
```{r}
# find parameters of the distribution of percent organic matter for each species
# aleuticum
fitOM.a <- fitdistr(serpA$Ompct,"gamma")
print(c(fitOM.a$estimate[1],fitOM.a$estimate[2]))

#viridimontanum
fitOM.v <- fitdistr(serpV$Ompct,"gamma")
print(c(fitOM.v$estimate[1],fitOM.v$estimate[2]))

# create random data that amplifies the existing difference
# aleuticum
datBlingA <- rgamma(6,shape=15,rate=.3)
sp <- rep("A",length=6)
datBlingA <- data.frame(cbind(datBlingA,sp))

# viridimontanum
datBlingV <- rgamma(9,shape=1.5,rate=.15)
sp <- rep("V",length=9)
datBlingV <- data.frame(cbind(datBlingV,sp))

# Trying to combine this data into a single dataframe for the ANOV() test, but I can't figure out how to coerce it to that form.  
```





