---
title: "Homework 8 - Functions for Statistical Analysis in R"
author: "Morgan Southgate"
date: "March 8, 2017"
output: html_document
---


# Function: Linear Regression
### Description: Fits a regression line to a scatterplot of x and y values
### Input: continuous x & y variables
### Output: abline & scatterplot
```{r, echo=TRUE}
linReg <- function(xVar= 1:100,yVar=runif(100,min=0,max=100),
                   dataframe=data.frame(xVar,yVar)){
  linRegMod <- lm(yVar~xVar,data=dataframe)
  linRegOut <- c(slope=summary(linRegMod)$coefficients[2,1],
                 pValue=summary(linRegMod)$coefficients[2,4])
    return(linRegOut)}

linReg()
```

##Function for plotting results of linear regression
```{r, echo=TRUE}
linRegPlot <-  function(xVar=1:100,yVar=runif(100,min=0,max=100),dataframe=data.frame(xVar,yVar)){
  lrplot <- plot(y=dataframe$yVar,
                       x=dataframe$xVar,
                      cex=2, 
                      pch=21,
                      bg="purple",
                      main="Linear Regression",
                      xlab="X Variable", 
                      ylab="Y variable",
                      xlim=range(xVar),
                      ylim=range(yVar))
  linRegMod <- lm(yVar~xVar,data=dataframe)
  abline(linRegMod)
  return(lrplot)
}

linRegPlot()
```

##Testing linReg() & linRegPlot() with tiny data set
```{r, echo=TRUE}
# make tiny data set
rdat <- data.frame(xVar=runif(10,min=0,max=20),yVar=runif(10,min=20,max=40))

# test linReg function with rdat
linReg(rdat)

# test linRegPlot function with rdat
linRegPlot(rdat)
```

# Function: Logistic Regression
## Description: 
## Input: continuous x variable & categorical y variable
## Output: estimate of xvar and p value of xvar

```{r, echo=TRUE}
logReg <- function(xVar=rgamma(n=20,shape=5,scale=5), 
                   yVar=rbinom(n=20,size=1,p=0.5)){
  logRegMod <- glm(yVar~ xVar,
                   family=binomial(link="logit"))
  logRegOut <- c(xVarEst=summary(logRegMod)$coefficients[2,1],
                 pValue=summary(logRegMod)$coefficients[2,4])
   return(logRegOut)}

logReg()
```

## Function for plotting results of logistic regression analysis
```{r, echo=TRUE, message=FALSE, warning=FALSE}
logRegPlot <- function(xVar=rgamma(n=20,shape=5,scale=5), 
                       yVar=rbinom(n=20,size=1,p=0.5),dataframe = data.frame(xVar,yVar)){
lrResults <- glm(yVar ~ xVar, family="binomial"(link="logit"))
LRplot <- plot(dataframe$xVar, y=dataframe$yVar,
                   pch=21,
                   bg="tan",
                   cex=1,
                   main="Logistic Regression")
LRplot1 <- curve(predict(lrResults,
              data.frame(xVar=x),
              type="response"),
      add=TRUE,
      lwd=2)
return(LRplot1)}

logRegPlot()
```

## Test logReg and logRegPlot functions with tiny data set
```{r, echo=TRUE}
# tiny data set
xVar <- rgamma(n=30,shape=4,scale=4)
yVar <- rbinom(n=30,size=1,p=0.5)
df <- data.frame(xVar,yVar)
 
# test logReg
logReg(xVar=xVar,yVar=yVar)

# test logRegPlot
logRegPlot(xVar=xVar,yVar = yVar,dataframe=df)
```


# Function: ANOVA
## input: categorical x variable and continuous y variable
## output: p value
```{r, echo=TRUE}
analVar <- function(xVar=as.factor(rep(c("A","B","C","D","E"),each=3)),
                    yVar=c(rgamma(10,shape=5,scale=5),rgamma(5,shape=5,scale=10))){
                      df=data.frame(xVar,yVar)
          aovMod <- aov(yVar~xVar,data=df)
          aovOut <- summary(aovMod)[[1]][["Pr(>F)"]][1]
         return(aovOut)}
        
analVar()
```

## Function for plotting boxplot to represent ANOVA results
```{r, echo=TRUE}
analVarPlot <- function(xVar=as.factor(rep(c("A","B","C","D","E"),each=3)),yVar=c(rgamma(15,shape=5,scale=5))){
  df <- data.frame(xVar,yVar)
  aovMod <- aov(yVar~xVar,data=df)
  aovPlot <-  boxplot(yVar~xVar,
                      data=df,
                      col=c("red","yellow","blue","green","purple"),
                    xlab=names(xVar),ylab=names(yVar))
  return(aovPlot)}
  
analVarPlot() 
```

## test analVar and analVarPlot using tiny data set
```{r, echo=TRUE}
# make tiny data set
xVar1 <- as.factor(rep(c("a","b","c"),each=5))
yVar1 <- c(rgamma(10,shape=5,scale=5),rgamma(5,shape=5,scale=10))

# test analVar() using tiny data set
analVar(xVar=xVar1,yVar=yVar1)

# test analVarPlot() using tiny data set
analVarPlot(xVar=xVar1,yVar=yVar1)
```

# Function: Contingency Table
## Input: discrete independent variable and discrete dependent variable
## Output: 
```{r, echo=TRUE, message=FALSE, warning=FALSE}
contTable <- function(x=c(22,40,60),y=c(40,80,45),datamatrix=rbind(x,y)){
  rownames(datamatrix)=c("Cold","Warm")
  colnames(datamatrix)=c("Species1","Species2","Species3")
  contTableMod <- chisq.test(datamatrix)
  contTableOut <- print(chisq.test(datamatrix)[3])
          return(contTableOut)}

contTable()
```

## Function for plotting mosaic plot of data complementary to contingency table results
```{r, echo=TRUE, message=FALSE, warning=FALSE}
contPlot <- function(x=c(22,40,60),y=c(40,80,45),datamatrix=rbind(x,y)){
  rownames(datamatrix)=c("Cold","Warm")
  colnames(datamatrix)=c("Species1","Species2","Species3")
  mplot <- mosaicplot(x=datamatrix,col=c("red","blue","yellow"),shade=F)
  return(mplot)}

contPlot()
```

## Test contingency table functions using tiny data set
```{r, message=FALSE, warning=FALSE}
x <- c(2,10,7)
y <- c(20,40,50)
dm <- rbind(x,y)

# test contTable()
contTable(x,y,dm)

# test contPlot()
contPlot(x,y,dm)
```

